---
title: "Stats Exam 2016"
author: "Carmen Roberts"
date: "2016/08/30"
output: github_document
---

## Question 1

**Null Hypothesis:** There is no relationship between oral body temperature and heart rate between males and females.

**Alternative Hypothesis:** There is a relationsship between oral body temperature and heart rate between males and females.

**Statistical test(s):** Body temperature and heart rate are normally distributed, parametric, paired data, categorical data and is on an interval scale.Therefore, a one-way repeated-measures ANOVA will be used to test the hypothesis. Significance is set at p<0.05 

**Assumptions:** There is an equal variance across the groups. The errors are independent. The data is matching effective.


```{r, question_1, echo=FALSE, warning=FALSE}

#indicate which package(s) to use
library(readr)              
library(tidyr)
library(dplyr)

#import dataset using URL
q1 <- read_csv('https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2016/question1.csv')

#plot data
male <- plot(q1$body_temperature, q1$male)
female <- plot(q1$body_temperature, q1$female)

#summary statistics of dataset
summary(q1)

#tidy data by gathering
g <- gather(q1, "Body Temperature", "Heart rate", 2:3, na.rm = FALSE)

#summary statistics of gathered data
summary(g)

#one-way repeted-measures ANOVA for males
aov_male <- aov(body_temperature ~ male, data = q1)
summary(aov_male)

#one-way repeated-measures ANOVA for females
aov_female <- aov(body_temperature ~ female, data = q1)
summary(aov_female)

#one-way repeated-measures ANOVA for HR
aov_HR <- aov(male ~ female, data = q1)
summary(aov_HR)

```

**Conclusion:** For data from male participants show F=2.31; p=0.133 and for female participants F=5.528; p=0.022.We reject our null hypothesis and conclude that there is a difference in oral body temperature en heart rate between males and females.      



## Question 2

**Null Hypothesis:** The direction of the first stumble and the final position relative to the line is not associated with sex or hand dominance.

**Alternative Hypothesis:** The direction of the first stumble and the final position relative to the line is associated with sex and hand dominance.

**Statistical test(s):** The data is paired and non-parametric. Therefore, a Wilcoxon signed-rank test will be used to test the hypothesis.Significance is set at p<0.05. 

**Assumptions for non-parametric data:**  The errors are independent. The data matching is effective. Samples are drawn from the same shape distributions.


```{r, question_2, echo=FALSE, warning=FALSE} 

#indicate which package(s) to use
library(readr)
library(tidyr)
library(dplyr)

#import dataset using URL
q2 <- read_csv('https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2016/question2.csv')

#plot the data
plot(q2)

#summary statistic of dataset
summary(q2)

#Wilcoxon signed-rank test to test hypothesis
wilcox.test(q2$sex, q2$first_stumble, paired = FALSE)
wilcox.test(q2$sex, q2$final_position, paired = FALSE)
wilcox.test(q2$handedness, q2$first_stumble, paired = FALSE)
wilcox.test(q2$handedness, q2$final_position, paired = FALSE)
```

**Conclusions:** There is no significant relationship between gender and the direction of the first stumble with  p=0.421. There is a significant relationship between handedness and and the first stumble with p<2.2e-16.
Therefore, we reject the null hypothesis.

##Question 3

**Null Hypothesis:** There is no relationship between the run number, he running time and calories consumed.

**Alternative Hypothesis:** There is a relationship between the run number, the running time and the calories consumed. 

**Statistical test(s):** The data is paired and non-parametric. Therefore a Wilcoxon signed-rank test will be used to test the hypothesis.

**Assumptions:**

```{r, question_3, echo=FALSE}

#indicate which package(s) to use
library(readr)

#import dataset using URL
q3 <- read_csv('https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2016/question3.csv')

#plot the data
plot(x = time, y = calories)

#add linear regression line
q3_reg <- lm(time ~ calories, data = q3)

#QQ-plot to test for normality of residuals
qqnorm(q3_reg$residuals)
qqline(q3_reg$residuals)

#fitted vs residual plot to test for homoskedasticity
plot(q3_reg$fitted, q3_reg$residuals)

```

**Concluions:**


## Question 4

```{r, question_4, echo=FALSE}
foo <- rnorm(10000, mean = 60, sd = 3) # final mark
bar <- rnorm(10000, mean = 68, sd = 5) # project mark
baz <- rep(seq(from = 1997, to = 2006), each = 1) # years

year <- sample(baz, 150, replace = TRUE,
               prob = c(0.05, 0.05, 0.08, 0.08, 
                          0.1, 0.13, 0.14, 0.13, 0.12, 0.12))
project_mark <- sample(bar, 150, replace = TRUE)
final_mark <- sample(foo, 150, replace = TRUE)

plot_data <- data_frame(year = year,
                        project_mark = round(project_mark, 1),
                        final_mark = round(final_mark, 1)) %>%
    arrange(year)

#plot the data

```
    











